---
title: "R Programming Assignment 1"
author: "Millie"
date: "Friday, January 22, 2016"
output: html_document
---

##Set directory- specdata
For this programming assignment you will need to unzip this file and create the directory 'specdata'. Once you have unzipped the zip file, do not make any modifications to the files in the 'specdata' directory. In each file you'll notice that there are many days where either sulfate or nitrate (or both) are missing (coded as NA). This is common with air pollution monitoring data in the United States.
```{r} 
setwd("C:/Users/User/Desktop/Coursera/R Programming/specdata")
F001= read.csv(file = "/Users/User/Desktop/Coursera/R Programming/specdata/001.csv", sep=",")
getwd()
```


##Part 1
Write a function named 'pollutantmean' that calculates the mean of a pollutant (sulfate or nitrate) across a specified list of monitors. The function 'pollutantmean' takes three arguments: 'directory', 'pollutant', and 'id'. Given a vector monitor ID numbers, 'pollutantmean' reads that monitors' particulate matter data from the directory specified in the 'directory' argument and returns the mean of the pollutant across all of the monitors, ignoring any missing values coded as NA. A prototype of the function is as follows
```{r}
pollutantmean= function(id=1:332){
#from-http://stackoverflow.com/questions/3397885/how-do-you-read-in-multiple-txt-files-into-r
    filenames <- list.files(pattern="*.csv", full.names=TRUE)
    ldf <- lapply(filenames, read.csv)
    samples=as.data.frame(do.call("rbind", ldf[id]))
  #'directory' is a character vector of length 1 indicating the location of the CSV files
#directory=specdata
  # 'pollutant' is a chartacter vector of length 2 indicating the names of the pllutant for which we will calculate the mean; either "sulfate" or "nitrate".
   #pollutant=c("sulfate", "nitrate")
  #'id' is an integer vector indicatinf the monitor ID numbers to be used
  #Return the mean of the pollutant across all monitors list in the 'id' vector (ignoring NA values)

apply(samples[,2:3], 2, function(x) mean(x, na.rm=TRUE))
#if(pollutant=="sulfate"){
#        print(meanpol[1])
#} else {
#        print(meanpol[2])
#} 


#if(pollutant=="sulfate"){
        #mean(samples$sulfate, na.rm = TRUE) 
        #}if else(pollutant=="nitrate"){ 
               # mean(samples$nitrate, na.rm = TRUE)
               # }
  
  #NOTE: Do not round the result!
}
```

Something like:
source("pollutantmean.R")
pollutantmean("specdata", "sulfate", 1:10)
 [1] 4.064
pollutantmean("specdata", "nitrate", 70:72)
 [1] 1.706
pollutantmean("specdata", "nitrate", 23)
 [1] 1.281

##Part 2
Write a function that reads a directory full of files and reports the number of completely observed cases in each data file. The function should return a data frame where the first column is the name of the file and the second column is the number of complete cases. A prototype of this function follows
```{r}
complete= function(id=1:332){
  #from-http://stackoverflow.com/questions/3397885/how-do-you-read-in-multiple-txt-files-into-r
  # id=1:3
filenames <- list.files(pattern="*.csv", full.names=TRUE)
ldf <- lapply(filenames, read.csv)
samples=as.data.frame(do.call("rbind", ldf[id]))

NaRowSum=rowSums(is.na(samples[,2:3])) 
final=as.data.frame(cbind(samples, NaRowSum))
print(table(samples$ID, NaRowSum==0))
  #'directory' is a character vector of length 1 indicating the loation of the CSV files
  #'id' is an integer vector indicating the monitor ID numbers to be used
  #Return a data frame of the form:
  #id nobs
  #1  117
  #2  1041
  #...
  #where 'id' is the monitor ID number and 'nobs' is the number of complete cases
}
```

Something like:
source("complete.R")
complete("specdata", 1)
   id nobs
 1  1  117
complete("specdata", c(2, 4, 8, 10, 12))
   id nobs
 1  2 1041
 2  4  474
 3  8  192
 4 10  148
 5 12   96
complete("specdata", 30:25)
   id nobs
 1 30  932
 2 29  711
 3 28  475
 4 27  338
 5 26  586
 6 25  463
complete("specdata", 3)
   id nobs
 1  3  243

##Part 3
Write a function that takes a directory of data files and a threshold for complete cases and calculates the correlation between sulfate and nitrate for monitor locations where the number of completely observed cases (on all variables) is greater than the threshold. The function should return a vector of correlations for the monitors that meet the threshold requirement. If no monitors meet the threshold requirement, then the function should return a numeric vector of length 0. A prototype of this function follows
```{r}
source("complete.R")
corr= function(directory, threshold=0){
filenames <- list.files(pattern="*.csv", full.names=TRUE)
ldf <- lapply(filenames, read.csv)
samples=as.data.frame(do.call("rbind", ldf[id]))

samples[complete.cases(samples),]
        
#'directory' is a character vector of length 1 indicating the location of the CSV files
  #'threshold' is a numeric vector of length 1 indicating the number of completely observed observations (on all cariables) required to compute the correlation between nitrate and sulfate; the default is 0
  #Return a numeric vector of correlations
  #NOTE: Do not round the result!
}
```
For this function you will need to use the 'cor' function in R which calculates the correlation between two vectors. Please read the help page for this function via '?cor' and make sure that you know how to use it.

```{r}
#various subsetting methods: http://www.r-bloggers.com/say-it-in-r-with-by-apply-and-friends/
#apply and lists of data: https://www.datacamp.com/community/tutorials/r-tutorial-apply-family
#apply to import: http://brianmannmath.github.io/blog/2014/01/20/using-lapply-to-import-files-to-r/
threshold=50
filenames= list.files(pattern="*.csv", full.names=TRUE)
ldf= lapply(filenames, read.csv)
samples= as.data.frame(do.call("rbind", ldf))

complete= samples[complete.cases(samples),]
thresh=subset(complete, ID>threshold)

for(i in unique(complete)){
Cor=cor.test(F001[,2], F001[,3], use="complete.obs", method="pearson") 
Cor[4]
quantile(Cor)
}
```


Something like:
source("corr.R")
source("complete.R")
cr <- corr("specdata", 150)
head(cr)
 [1] -0.01896 -0.14051 -0.04390 -0.06816 -0.12351 -0.07589
summary(cr)
    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
 -0.2110 -0.0500  0.0946  0.1250  0.2680  0.7630
cr <- corr("specdata", 400)
head(cr)
 [1] -0.01896 -0.04390 -0.06816 -0.07589  0.76313 -0.15783
summary(cr)
    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
 -0.1760 -0.0311  0.1000  0.1400  0.2680  0.7630
cr <- corr("specdata", 5000)
summary(cr)
    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
 
length(cr)
 [1] 0
cr <- corr("specdata")
summary(cr)
    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
 -1.0000 -0.0528  0.1070  0.1370  0.2780  1.0000
length(cr)
 [1] 323
